---
title: "Final - 3"
author: "Sowmya Maddali"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
pacman::p_load(fredr, ggplot2, tidyverse, tsbox, tseries,
               tsibble, forecast, tseries, lubridate,
               expsmooth, WDI, knitr, broom, ggpubr,
               stargazer, urca, patchwork, ForecastComb, vars)
```



# Combining the dataframes

```{r}
comb_df <- data.frame(cpi$date, cpi$value, ppi$value, ppi_fg$value, ppi_fcf$value)
head(comb_df)
```



# Checking for co-integration

```{r}
johansen_test <- ca.jo(comb_df[, c('cpi.value', 'ppi.value', 'ppi_fg.value', 'ppi_fcf.value')],
                       type = 'eigen', spec = 'transitory',
                       ecdet = 'none', K = 2)

summary(johansen_test)
```


# Visualizing the linear combination

```{r}
comb_df <- comb_df %>%
  mutate(q = cpi.value + 0.5844253 * ppi.value -5.6989774 * ppi_fg.value + 3.4875670 * ppi_fcf.value)

ggplot(comb_df) + theme_bw(base_size = 12) +
  geom_line(aes(x = cpi.date, y = q), color = "blue") +
  ggtitle("A Linear Combination of all 4 variables") +
  xlab("Date") + ylab("Linear Combination Value")
```


# Optimal lag length

```{r}
lag <- VARselect(comb_df[2:4])
lag
```


# Taking the difference of all the series and combining it in 1 dataframe


```{r}
comb_df_diff <- data.frame(
  cpi_df = diff(comb_df[, 2]),
  ppi_df = diff(comb_df[, 3]),
  ppifg_df = diff(comb_df[, 4]),
  ppifcf_df = diff(comb_df[, 5])
)

head(comb_df_diff)
```


# Train and test set

```{r}
# train set
train_df <- comb_df_diff[1:241,]
head(train_df)

# test set
test_df <- comb_df_diff[242:251,]
head(test_df)
```


# Constructing the VAR model

```{r}
# model - 1
var_1 <- VAR(comb_df_diff, p = 6, type = 'none', season = 12)
summary(var_1)
```



```{r}
# model - 2
alt_1 <- VAR(comb_df_diff, p = 2, type ='none')
summary(alt_1)
```

```{r}
# model - 3
alt_2 <- VAR(comb_df_diff, p = 1, type = 'const')
summary(alt_2)
```


# Forecasting

```{r}
pred_1 <- predict(var_1, n.ahead = 10)
pred_2 <- predict(alt_1, n.ahead = 10)
pred_3 <- predict(alt_2, n.ahead = 10)
```


```{r}
print(pred_1$fcst$cpi_df[,1])
print(pred_1$fcst$ppi_df[,2])
print(pred_1$fcst$ppifg_df[,3])
print(pred_1$fcst$ppifcf_df[,4])
```


# Plotting the forecasts - Model 1

```{r}
# Plot - 1

plot(test_df$cpi_df, type = "l", col = "blue",ylim=c(-0.8,1),
     main = "Actual vs Predicted CPI values",
     xlab = "Time Period",
     ylab = "Test CPI values")

lines(pred_1$fcst$cpi_df[,1], col = "red")

legend("bottomleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 2

plot(test_df$ppi_df, type = "l", col = "blue",ylim=c(-3,1.9),
     main = "Actual vs Predicted PPI values",
     xlab = "Time Period",
     ylab = "Test PPI values")

lines(pred_1$fcst$ppi_df[,1], col = "red")

legend("bottomleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 3

plot(test_df$ppifg_df, type = "l", col = "blue",ylim=c(-2,3.3),
     main = "Actual vs Predicted PPIFG values",
     xlab = "Time Period",
     ylab = "Test PPIFG values")

lines(pred_1$fcst$ppifg_df[,1], col = "red")

legend("topleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 4

plot(test_df$ppifcf_df, type = "l", col = "blue",ylim=c(-3,5),
     main = "Actual vs Predicted PPIFCF values",
     xlab = "Time Period",
     ylab = "Test PPIFCF values")

lines(pred_1$fcst$ppifcf_df[,1], col = "red")

legend("topleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)
```

# Plotting the forecasts - Model 2

```{r}
# Plot - 1

plot(test_df$cpi_df, type = "l", col = "blue",ylim=c(-0.8,1),
     main = "Actual vs Predicted CPI values",
     xlab = "Time Period",
     ylab = "Test CPI values")

lines(pred_2$fcst$cpi_df[,1], col = "red")

legend("bottomleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 2

plot(test_df$ppi_df, type = "l", col = "blue",ylim=c(-3,1.9),
     main = "Actual vs Predicted PPI values",
     xlab = "Time Period",
     ylab = "Test PPI values")

lines(pred_2$fcst$ppi_df[,1], col = "red")

legend("bottomleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 3

plot(test_df$ppifg_df, type = "l", col = "blue",ylim=c(-2,3.3),
     main = "Actual vs Predicted PPIFG values",
     xlab = "Time Period",
     ylab = "Test PPIFG values")

lines(pred_2$fcst$ppifg_df[,1], col = "red")

legend("topleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 4

plot(test_df$ppifcf_df, type = "l", col = "blue",ylim=c(-3,5),
     main = "Actual vs Predicted PPIFCF values",
     xlab = "Time Period",
     ylab = "Test PPIFCF values")

lines(pred_2$fcst$ppifcf_df[,1], col = "red")

legend("topleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)
```


# Plotting the forecasts - Model 3

```{r}
# Plot - 1

plot(test_df$cpi_df, type = "l", col = "blue",ylim=c(-0.8,1.2),
     main = "Actual vs Predicted CPI values",
     xlab = "Time Period",
     ylab = "Test CPI values")

lines(pred_3$fcst$cpi_df[,1], col = "red")

legend("bottomleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 2

plot(test_df$ppi_df, type = "l", col = "blue",ylim=c(-3,1.9),
     main = "Actual vs Predicted PPI values",
     xlab = "Time Period",
     ylab = "Test PPI values")

lines(pred_3$fcst$ppi_df[,1], col = "red")

legend("bottomleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 3

plot(test_df$ppifg_df, type = "l", col = "blue",ylim=c(-3,3.3),
     main = "Actual vs Predicted PPIFG values",
     xlab = "Time Period",
     ylab = "Test PPIFG values")

lines(pred_3$fcst$ppifg_df[,1], col = "red")

legend("bottomleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)

# Plot - 4

plot(test_df$ppifcf_df, type = "l", col = "blue",ylim=c(-3,5),
     main = "Actual vs Predicted PPIFCF values",
     xlab = "Time Period",
     ylab = "Test PPIFCF values")

lines(pred_3$fcst$ppifcf_df[,1], col = "red")

legend("topleft", legend = c("Actual", "Predicted"),
       col = c("blue", "red"), lty = 1)
```



# Loss functions - Model 1

```{r, warning=FALSE}
loss_1 <- loss_functions(pred_1$fcst$cpi_df[,1], test_df$cpi_df)
print(loss_1[1:2])

loss_2 <- loss_functions(pred_1$fcst$ppi_df[,1], test_df$ppi_df)
print(loss_2[1:2])

loss_3 <- loss_functions(pred_1$fcst$ppifg_df[,1], test_df$ppifg_df)
print(loss_3[1:2])

loss_4 <- loss_functions(pred_1$fcst$ppifcf_df[,1], test_df$ppifcf_df)
print(loss_4[1:2])
```


# Loss functions - Model 2

```{r, warning=FALSE}
loss_1 <- loss_functions(pred_2$fcst$cpi_df[,1], test_df$cpi_df)
print(loss_1[1:2])

loss_2 <- loss_functions(pred_2$fcst$ppi_df[,1], test_df$ppi_df)
print(loss_2[1:2])

loss_3 <- loss_functions(pred_2$fcst$ppifg_df[,1], test_df$ppifg_df)
print(loss_3[1:2])

loss_4 <- loss_functions(pred_2$fcst$ppifcf_df[,1], test_df$ppifcf_df)
print(loss_4[1:2])
```


# Loss functions - Model 3

```{r, warning=FALSE}
loss_1 <- loss_functions(pred_3$fcst$cpi_df[,1], test_df$cpi_df)
print(loss_1[1:2])

loss_2 <- loss_functions(pred_3$fcst$ppi_df[,1], test_df$ppi_df)
print(loss_2[1:2])

loss_3 <- loss_functions(pred_3$fcst$ppifg_df[,1], test_df$ppifg_df)
print(loss_3[1:2])

loss_4 <- loss_functions(pred_3$fcst$ppifcf_df[,1], test_df$ppifcf_df)
print(loss_4[1:2])
```


# IRF plots

```{r}
irf_1 <- irf(var_1, impulse = "cpi_df", 
             response = c("cpi_df", "ppi_df", "ppifg_df", "ppifcf_df"),
             ortho = T,
             n.ahead = 10)
plot(irf_1)

```


```{r}
irf_2 <- irf(var_1, impulse = c("cpi_df", "ppi_df", "ppifg_df", "ppifcf_df"), 
             response = "cpi_df",
             ortho = T,
             n.ahead = 10)
plot(irf_2)

```


# Error decomposition 

```{r}
fevd_vars <- fevd(var_1, n.ahead = 10)
kable(fevd_vars$cpi_df, digits = 3)
kable(fevd_vars$ppi_df, digits = 3)
kable(fevd_vars$ppifg_df, digits = 3)
kable(fevd_vars$ppifcf_df, digits = 3)
```


```{r}
irf_3 <- irf(var_1, n.ahead = 10)
plot(irf_3)

```






