---
title: "Midterm - (Producer Price Index by Commodity: Final Demand: Finished Consumer Foods (WPSFD4111))"
author: "Sowmya Maddali"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
pacman::p_load(fredr, ggplot2, tidyverse, tsbox, tseries,
               tsibble, forecast, tseries, lubridate,
               expsmooth, WDI, knitr, broom, ggpubr,
               stargazer, urca, patchwork, ForecastComb)
```



```{r}
fredr_set_key("5a3f5a4d628dad0eeebb7dc5711f691c")
ppi_fcf <- fredr(series_id = "WPSFD4111",
                    observation_start = as.Date("1999-01-01"),
                    observation_end = as.Date("2019-12-01")) %>%
  select(date, series_id, value) %>%
  mutate(index_first_diff = value - lag(value),
         index_second_diff = difference(value, differences = 2)) %>%
  tsibble(index = date)
```



# Plot

```{r}
plot(ppi_fcf$date, ppi_fcf$value, xlab = "Time", ylab = "PPI Finished Consumer Foods values",
    main = "PPI: Final Demand: Finished Consumer Foods")
```



# Summary

```{r}
summary(ppi_fcf)
```



# Intord results

```{r}
intord(ppi_fcf$value, year = 1999, freq = 12, period = 1)

# it is trend stationary. Here trend needs to be removed.
```

Level series is not stationary but the first difference is stationary


# ACF and PACF

```{r}
ggAcf(ppi_fcf$index_first_diff) + ggtitle("ACF plot for PPI Consumer Foods") | ggPacf(ppi_fcf$index_first_diff) + ggtitle("PACF plot for PPI Consumer Foods")
```


ACF & PACF decay at the same rate, so ARMA models will be used.


# Train and test split

```{r}
train_ppi_fcf <- ppi_fcf$value[1:246]
test_ppi_fcf <- ppi_fcf$value[247:252]

ts_train_ppifcf <- ts(train_ppi_fcf, start = c(1999, 1), freq = 12)
ts_test_ppifcf <- ts(test_ppi_fcf, start = c(2019, 7), freq = 12)

glimpse(ts_train_ppifcf)
tail(ts_test_ppifcf)
```


# Modeling

```{r}
# Model - 1
# p = 5, d = 1, q = 4

ppifcf_model_1 <- Arima(ts_train_ppifcf,
                      order = c(5, 1, 4))

print(ppifcf_model_1)

```


```{r}
# Model - 2
# p = 4, d = 1, q = 3

ppifcf_model_2 <- Arima(ts_train_ppifcf,
                      order = c(4, 1, 4))

print(ppifcf_model_2)

```


```{r}
# Model - 3
# p = 2, d = 1, q = 2

ppifcf_model_3 <- Arima(ts_train_ppifcf,
                      order = c(2, 1, 2))

print(ppifcf_model_3)

```

Model 2 is the best model according to AIC values.

# Brute force model

```{r}
ppifcf_brute_force_model <- auto.arima(ts(ts_train_ppifcf),
                                ic = 'aic',
                                trace = T)
```

Model 2 has the lowest AIC value


# Serial Correlation

```{r}
ggAcf(ppifcf_model_1$residuals) + ggtitle("Serial Correlation for Model 1")
ggAcf(ppifcf_model_2$residuals) + ggtitle("Serial Correlation for Model 2")
ggAcf(ppifcf_model_3$residuals) + ggtitle("Serial Correlation for Model 3")
```



# Ljung-Box Q Statistic in R

```{r}
Box.test(ppifcf_model_1$residuals, 
         type = "Ljung-Box")

Box.test(ppifcf_model_2$residuals, 
         type = "Ljung-Box")

Box.test(ppifcf_model_3$residuals, 
         type = "Ljung-Box")
```

All the p-values are above 0.05, hence there is no serial correlation being exhibited here.



# Forecast

```{r}
# Model - 1
ppifcf_forecast_model_1 <- forecast(ppifcf_model_1, 
                                 h = 6)
print(ppifcf_forecast_model_1)


# Model - 2
ppifcf_forecast_model_2 <- forecast(ppifcf_model_2, 
                                 h = 6)
print(ppifcf_forecast_model_2)


# Model - 3
ppifcf_forecast_model_3 <- forecast(ppifcf_model_3,
                                 h = 6)
print(ppifcf_forecast_model_3)

```


# Autoplot

```{r}
autoplot(ppifcf_forecast_model_1, xlim = c(2015, 2020), ylim = c(200,220))
autoplot(ppifcf_forecast_model_2, xlim = c(2015, 2020), ylim = c(200,220))
autoplot(ppifcf_forecast_model_3, xlim = c(2015, 2020), ylim = c(200,220))
```


# Loss Functions

```{r, warning=FALSE}
print("Model 1")
# Model - 1
ppifcf_loss_model_1 <- loss_functions(ppifcf_forecast_model_1$mean, ts_test_ppifcf)
print(ppifcf_loss_model_1[3:4])

print("Model 2")
# Model - 2
ppifcf_loss_model_2 <- loss_functions(ppifcf_forecast_model_2$mean, ts_test_ppifcf)
print(ppifcf_loss_model_2[3:4])

print("Model 3")
# Model - 3
ppifcf_loss_model_3 <- loss_functions(ppifcf_forecast_model_3$mean, ts_test_ppifcf)
print(ppifcf_loss_model_3[3:4])
```

Model 2 is the best model according to AIC and loss functions


# Granger-Bates

```{r}
ppifcf_combination_object <- foreccomb(test_ppi_fcf,
                                cbind(ppifcf_forecast_model_1$mean[1:6],
                                      ppifcf_forecast_model_2$mean[1:6],
                                      ppifcf_forecast_model_3$mean[1:6]))
print(ppifcf_combination_object)

ppifcf_granger_bates <- comb_BG(ppifcf_combination_object)
print(ppifcf_granger_bates)
```




```{r}
# Model forecasts (from part iv)
model1_forecast <- c(210.2947, 210.7514, 210.7123, 211.0236, 211.0116, 211.2637) 
model2_forecast <- c(210.6057, 210.2897, 210.2661, 210.7847, 211.4159, 211.5679)
model3_forecast <- c(210.1603, 210.4701, 210.4623, 210.3750, 210.3989, 210.4181)

# Actual test observations
test_obs <- c(211.0, 209.7, 210.6, 212.7, 216.3, 216.2)  

# Assume model 1 is preferred 
preferred_model <- model2_forecast

# Create forecast combination object
fc_object <- foreccomb(test_obs, cbind(model1_forecast, model2_forecast, model3_forecast))

# Get Granger-Bates combined forecast
gb_forecast <- comb_BG(fc_object)

# Plot forecasts
plot(test_obs, type="o", col="black", ylim=c(209,217),
     xlab="Time", ylab="Value")
lines(preferred_model, col="blue", lwd=2) 
lines(gb_forecast$Fitted, col="red", lwd=2)
legend("topleft", legend=c("Test Observations", "Forecast Model 2", "Granger-Bates Combination Model"),
       col=c("black", "blue", "red"), lwd=c(1,2,2), cex=0.8)
title("Forecast Accuracy Comparison")
```



```{r}
# Compute forecast errors
preferred_error <- sqrt(mean((test_obs - preferred_model)^2)) 
gb_error <- sqrt(mean((test_obs - gb_forecast$Fitted)^2))

# Print forecast errors
print(paste("Preferred model RMSE:", round(preferred_error, 2)))
print(paste("Granger-Bates RMSE:", round(gb_error, 2)))
```




